---
title: "Quarto/Markdown example"
format: 
  html: 
   toc: true
   self-contained: true
   df-print: paged
   code-link: true
lightbox: true
---

```{r setup}
#| include: FALSE
knitr::opts_chunk$set(echo = TRUE, 
                      message=FALSE, 
                      warning =FALSE)
```

Paste all this stuff into a new empty Quarto document and knit it to create some HTML output.

## Example

Here's an example of loading, filtering, and recoding a variable from the [NHANES](https://www.cdc.gov/nchs/nhanes/index.html) data, and then using `gt` to produce a formatted table. Note that the `code-fold` option is set to `true` for both chunks, so the R code is collapsed until you press the button to expand it:

```{r}
#| code-fold: true
library(tidyverse)
library(gt)
nhanes<-read_csv("https://raw.githubusercontent.com/Neilblund/SDSB340-Spring-2026/refs/heads/master/Data/nhanes.csv")

comp_hours<-nhanes|>
  filter(Age>=18)|>
  mutate(Education = fct_relevel(Education, '8th Grade', 
                                 '9 - 11th Grade',
                                 'High School',
                                 'Some College',
                                 'College Grad'
  ))|>
  group_by(Education)|>
  summarize( CompHrsDay= mean(CompHrsDay%in%c('4_hr', 'More_4_hr'), 
                              na.rm=T))|>
  
  drop_na()

comp_hours|>
  gt()|>
  fmt_percent(columns = CompHrsDay)|>
  cols_label(CompHrsDay = "% 4+ hours on the computer daily")|>
  tab_source_note("Source: National Health and Nutrition Examination Survey 2024")


```


Here's an example of creating a figure in ggplot to display the same data.


```{r}
#| code-fold: true
comp_hours|>
  ggplot(aes(x=Education, y=CompHrsDay)) +
  geom_bar(stat='identity', fill='maroon') +
  theme_minimal() +
  labs(y="Percent",
       caption ="Source: National Health and Nutrition Examination Survey 2024",
       title = 'Respondents with 4+ hours on the computer daily',
       subtitle='by education'
       ) +
  scale_y_continuous(labels = scales::percent) +
  theme(
    axis.text.x = element_text(size=10),
    axis.text.y = element_text(size=10),
    axis.title =element_text(size=12, face="bold"),
    title = element_text(size=14, face='bold')
)
```

